Kalman tracking and estimation of continuous gravitational waves with a pulsar timing array

### 2023-08-21 docs/manuscript/paper.tex
bayes, 2000 points

The parameters which govern the deterministic evolution of the pulsar spin frequency, $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$ are well-determined by existing radio timing observations. We can identify $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$ with the pulsar  barycentric rotation frequency and the time derivative of barycentric rotation frequency quoted by pulsar observation catalogues. For the 12.5 year NANOGrav pulsars the median error in  the barycentric rotation frequency is $\sim 7 \times 10^{-13}$ Hz, and in the time derivative of barycentric rotation frequency is $\sim 1.8 \times 10^{-20}$ s $^{-2}$. Whilst it is then clear that $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$ are typically measured with very high precisions, for this work we adopt a broader uniform prior, $\pm 10 \%$ either side of the true value. In this way we can test how well our method performs without requiring exceptionally precise and accurate measurements of the pulsar parameters to be made \textit{a priori}. Instead we consider the pulsar parameters similarly to the GW source parameters and estimate them optimally within a consistent framework. Adopting looser priors in this way allows us to subject our method to a more stringent test and confirm that we still estimate GW parameters without highly accurate initial estimates of $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$. \newline

The pulsar distances $d^{(n)}$ are typically not as well constrained as $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$, with uncertainties typically on the order of $\sim 10 \%$ \citep{Arzoumanian2018ApJS..235...37A,Yao2017}. However for this work, whilst the pulsar distance is used when generating synthetic data, as discussed in Section \ref{sec:parameter_estim} the pulsar distance is not used for our inference model and so we do not need to set a prior on $d$. Similarly we do not set a prior on $\gamma^{(n)}$; since $[\gamma^{(n)}]^{-1} >> T_{\rm obs}$ this means that $\gamma^{(n)}$ is effectively ``unobservable" over the decadal timescales that we are interested in. That is, for $T_{\rm obs}$ =10 years the solution of Equation \eqref{eq:spinevol} is effectively independent of the choice of $\gamma^{(n)}$ as long as the condition $[\gamma^{(n)}]^{-1} \gg T_{\rm obs}$ is satisfied. It is therefore sufficient to consider $\gamma^{(n)}$ to be known \textit{a priori} and set it at its true injected value. We briefly explored setting an uninformative prior on $\gamma$ over e.g. LogUniform($10^{-15}, 10^{-10}$) as well as setting $\gamma^{(n)}$ at some fixed value away from the true injected value (e.g. set $\gamma^{(n)} = 10^{-14} \text{s}^{-1}$ rather than $10^{-13} \text{s}^{-1}$) but the results are unchanged. The majority of pulsars in our synthetic PTA have $\sigma^{(n)} \sim 10^{-20} - 10^{-22}$ as calculated from Equations \eqref{eq:sigmap} and \eqref{eq:sigmap_f}. For these pulsars we set an uninformative broad prior of LogUniform($10^{-23}, 10^{-19}$). The single exception is PSR J1939+2134 which has a particularly large $\dot{f}_{\rm p}^{(n)}$ compared to the other pulsars in the array and so $\sigma \sim 10^{-16}$. For our artificial dataset we set $\sigma = 10^{21}$ for this pulsar in order to bring the process noise for all pulsars into a consistent range that can be described by a constrained prior

Since we are not setting priors on $\gamma$ or $d$ this reduces the dimensionality of the parameter space to $7 + 3N$. We use the notation $\boldsymbol{\theta}_{\rm psr, reduced}$ to refer to the reduced parameter space. Explicitly, 
\begin{equation}
	\boldsymbol{\theta}_{\rm psr, reduced} = \left \{ f_{\rm em}^{(n)}(t_1),\dot{f}_{\rm em}^{(n)}(t_1),\sigma^{(n)}\right\}_{1\leq n \leq N} \ .
\end{equation}
All the injected static parameters and their corresponding priors for this representative example are summarised in Table \ref{tab:parameters_and_priors}.
\begin{table}
	\centering
	\resizebox{\columnwidth}{!}{%
	\renewcommand{\arraystretch}{1.5} % Default value: 1
	\begin{tabular}{lccll}
		\toprule
		&Parameter & Injected Values & Units & Prior  \\
		\hline
		\multirow{7}{2mm}{$\boldsymbol{\theta}_{\rm gw}$} & $\Omega$       & $5 \times 10^{-7}$ & Hz & LogUniform($10^{-9}$, $10^{-5}$) \\
	  & $\alpha$          & $1.0$  & radians & Uniform($0, 2 \pi $)\\
	  & $\delta$              & $1.0$  & radians & Uniform($-\pi/2, \pi/2$) \\
	  & $\psi$              & $2.50$ & radians & Uniform($0, 2 \pi $) \\
	  & $\Phi_0$          & $0.20$ & radians & Uniform($0, 2 \pi $) \\
	  & $h_0$            & $10^{-12}$ & - & LogUniform($10^{-15}$, $10^{-9}$) \\
	  & $\iota$             & $1.0$ & radians & Uniform($0, \pi$) \\ 
		\hline
		\multirow{5}{2mm}{$\boldsymbol{\theta}_{\rm psr}$} & $f_{\rm em}^{(n)} (t_1)$       & $f_{\rm ATNF}^{(n)}$ & Hz & Uniform(0.9$f_{\rm ATNF}^{n}$, $1.1 f_{\rm ATNF}^{n}$) \\
		& $\dot{f}_{\rm em}^{(n)} (t_1)$       & $\dot{f}_{\rm ATNF}^{(n)}$ & s$^{-2}$ & Uniform(0.9$\dot{f}_{\rm ATNF}^{n}$, $1.1 \dot{f}_{\rm ATNF}^{n}$) \\
		&  $d^{(n)}$       &$d_{\rm ATNF}^{(n)}$  & m & - \\
		& $\sigma^{(n)}$              & $\sigma_{\rm sc}^{(n)}$ & $s^{-5/2}$ & LogUniform($10^{-23}, 10^{-19}$) \\
		& $\gamma$              & $10^{-13}$ & s$^{-1}$ & - \\
		\bottomrule
	\end{tabular}
}
	\caption{Summary of injected static parameters used for generating synthetic data in the representative example of Section \ref{sec:rep_example}, along with the choice of prior used for Bayesian inference on each parameter. The subscript ``ATNF" denotes values which have been obtained from the ATNF pulsar catalogue as described in Section \ref{sec:synt_pta}. The subscript ``sc" indicates that the injected value has been calculated using the empirical timing model, Equations \eqref{eq:sigmap}, \eqref{eq:sigmap_f} from \protect \cite{Shannon2010}.}
	\label{tab:parameters_and_priors}
\end{table}

@@

\begin{figure*}
	\includegraphics[width=0.8\textwidth, height =0.8\textwidth ]{images/representative_example_v2_GW}
	\caption{Posterior distribution for the GW source parameters $\boldsymbol{\theta}_{\rm gw}$ for a single realisation of the system noise. The vertical orange lines indicate the true injected values, c.f. Table \ref{tab:parameters_and_priors}. The contours in the 2D histograms denote the (0.5, 1, 1.5, 2)-$\sigma$ levels. We are able to accurately estimate each parameter of interest. $\Omega_{\rm nHz} = \Omega \times 10^{9}$ Hz and $h_{0, \times 10^12} = h_0 \times 10^{12}$.}
	\label{fig:corner_plot_1}
\end{figure*}

@@

\subsubsection{Detection} \label{sec:detection}
We frame the problem of claiming a detection of a GW in the noisy data in terms of a Bayesian model selection procedure as described in Section \ref{sec:model_selection}. $\mathcal{M}_1$ is the Earth-terms only model, i.e. Equation \eqref{eq:measuremen_earth}. The Bayes factor, $\beta$, as defined in Equation \eqref{eq:bayes} is presented in Fig \ref{fig:bayes} for our representative system, where we now vary the magnitude of the GW strain, $h_0$. The noise processes in the synthetic data are identical realisations for each value of $h_0$; the only change in the data is the change in $h_0$. We can see that there is a general log-linear relationship between $\beta$ and $h_0$ where the GW source is easily detectable with decisive evidence ($\beta \sim 100$) for $h_0 \gtrapprox 10^{-14}$. We take $\beta = 10$ as a tolerance cut-off above which we can accept $\mathcal{M}_1$ and claim a detection of the GW. For this system, given this realisation of the noise, the minimal detectable strain is $\sim 3 \times 10^{-15}$. This minimal detectable strain is of course particular to this system and will be influenced in general by e.g. $T_{\rm obs}$, $\sigma_{\rm m}$ as well as the parameters of the GW source such as its location on the sky. This will be explored in more detail in Section \ref{sec:parameter_space}. The small perturbations in the general log-linear $\beta - h_0$ trend arise due to noise in the nested sampling method. For these points the sampler converged sub-optimally for $\mathcal{M}_1$ and so the evidence $\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1)$ is lower than would be expected. This is purely a noise artefact; if one recreates the $\beta - h_0$ curve by rerunning the nested sampler these perturbations appear at different values of $h_0$, or else are not present at all. Increasing the number of live points $n_{\rm live}$ used by the nested sampler also dampens these perturbations and smooths the $\beta - h_0$ curve, since as discussed in Section \ref{sec:nested_sampling} the uncertainties in $\mathcal{Z}$ scale as $\mathcal{O}(1/ \sqrt{n_{\rm live}})$. Below $\beta = 10$ this noise in the sampler starts to dominate and the hierarchical relationship between $\mathcal{M}_0$ and  $\mathcal{M}_1$ fails. That is, the requirement that $\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1) > \mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_0)$ no longer holds. This is simply due to the fact that at these small strains the system is completely noise dominated and the two models $\mathcal{M}_1, \mathcal{M}_0$ cannot be separated.  \textcolor{red}{TK: Rather than discussing the bumps, maybe just run it again a few times and remove the bumps.}
\begin{figure}
	\includegraphics[width=\columnwidth]{images/PaperBayesRatioPlotn1000_stacked}
	\caption{Bayes ratio $\beta$ between the competing models $\mathcal{M}_1$ (GW present in data) and $\mathcal{M}_0$ (GW not present in data) at different GW strain magnitudes $,h_0$, for the representative example summarised in Table \ref{tab:parameters_and_priors}. The horizontal grey dashed line labels the detection tolerance cut-off of $\beta = 10$. The minimal detectable strain, below which $\beta < 10$, is $\sim 3 \times 10^{-15}$. \textcolor{red}{TK: probably need to regenerate this figure to get rid of the "bumps" which are just artefacts from a poor sampler convergence at that point}}
	\label{fig:bayes}
\end{figure}

\subsubsection{Multiple noise realisations} \label{sec:multiple_noise}
\begin{figure*}
	\includegraphics[width=0.8\textwidth, height =0.8\textwidth ]{images/stacked_GW_plot_1000}
	\caption{As Figure \ref{fig:corner_plot_1} but for 9 realisations of the noise processes. The inferred posterior distributions show strong agreement between the different noise realisations for the majority of static parameters. The exceptions are $\iota$ and $h$ which exhibit a larger degree of variance.} 
	\label{fig:corner_plot_2}
\end{figure*}
\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{images/compare_iotas_lines2}
	\caption{1D marginalized posteriors for $\iota$ over 9 realisations of the noise where $h$ is (a) an unknown free parameter (left panel) and (b) fixed at its true value $= 10^{-12}$. The left hand panel is equivalent to the marginalized posteriors for $\iota$ shown in Figure \ref{fig:corner_plot_2} (4th row, 4th column). The thick orange line labels the true injected value. Fixing $h$ reduces the variance between different noise realisations but results in a clear bias between the true and inferred values of $\iota$.}
	\label{fig:iota}
\end{figure*}
\begin{figure*}
	\setkeys{Gin}{width=\linewidth}

	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/distribution_omega_gw}
		\caption{$\Omega$}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/distribution_phi0_gw}
		\caption{$\Phi_0$}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/distribution_psi_gw}
		\caption{$\psi$}
	\end{subfigure}
	\medskip

	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/distribution_iota_gw}
		\caption{$\iota$}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/distribution_delta_gw}
		\caption{$\delta$}
	\end{subfigure}
	\hfill	
	\begin{subfigure}[b]{0.3\textwidth}
		\includegraphics[width=\textwidth]{images/distribution_alpha_gw}
		\caption{$\alpha$}
	\end{subfigure}
	\medskip

	\medskip
	\begin{subfigure}[b]{0.33\textwidth}
		\includegraphics{images/distribution_h}
		\caption{$h$}
	\end{subfigure}

	\caption{Distribution in the median values of the 1D marginalized posteriors inferred by nested sampling for each static parameters of $\boldsymbol{\theta}_{\rm gw}$ over 1000 noise realisations. The orange dashed line labels the true injection value and the red dashed line labels the median of the distribution. \textcolor{red}{TK: these plots are ugly. Is there a better way to show this?}}\label{fig:median_distriubutins}
\end{figure*}
The above discussion was for a single realisation of the noise processes. It is important to confirm that our method is robust against different realisations of the noise and that the specific realisation of the noisy data used in Sections \ref{sec:parameter_estim}, \ref{sec:detection} is not particularly advantageous or convenient for our method. To this end we take our representative example from Table \ref{tab:parameters_and_priors} and generate 1000 realisations of the process noise and measurement noise. For each realisation of the data we can then independently estimate the static parameters, $\boldsymbol{\theta}$. In Figure \ref{fig:corner_plot_2} we show the results for the  estimated values of $\boldsymbol{\theta}_{\rm gw}$ for 9 realisations of the noise. This figure is analogous to the single noise realisation of Figure \ref{fig:corner_plot_1}. We plot only 9 realisations of the noise rather than the full set of 1000 realisations so as to not overcrowd the figure. As in Section \ref{sec:parameter_estim} the estimates of $\boldsymbol{\theta}_{\rm psr}$ are recovered unambiguously across the 1000 noise realisations and for brevity we do not show the results here. \newline

@@

For each realisation of the noise we infer a marginalised 1-D posterior for each parameter. In turn, each posterior distribution has a median value. We plot the distribution of these medians over the 1000 noise realisations for each parameter of $\boldsymbol{\theta}_{\rm gw}$ in Figure \ref{fig:median_distriubutins}. For each of these distributions we can also calculate the median value (i.e. the median of the medians) and compare it with the true injection value. This is shown by the red and orange dashed lines respectively in the Figure. Analogous to the results we saw in Figure \ref{fig:corner_plot_2}, the distributions of the median values of the parameters $\Omega, \Phi_0, \psi, \delta$ and $\alpha$ are very narrow, with a generally small discrepancy between the inferred value and the injected value. Conversely, the distributions for $\iota$ and $h_0$ are much broader and exhibit a bias of $\sim 0.18$ radians and $0.17 \times 10^{-12}$ respectively. This agrees with the results we discussed for the 9 noise realisations in Figure \ref{fig:corner_plot_2} where a bias in $\iota$ and $h_0$ was also observed. In addition to  $\iota$ and $h_0$  exhibiting a strong bias, the parameters $\psi$ and $\alpha$ also exhibit a small bias of magnitude $\sim 0.1$ and $\sim 0.04$ radians respectively. This bias is again a result of dropping the pulsar terms from the measurement equation and will be discussed in Section \ref{sec:bias}.

@@

>>>>

The parameters which govern the deterministic evolution of the pulsar spin frequency, $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$ are well-determined by existing radio timing observations. We can identify $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$ with the pulsar  barycentric rotation frequency and the time derivative of barycentric rotation frequency quoted by pulsar observation catalogues. For the 12.5 year NANOGrav pulsars the median error in  the barycentric rotation frequency is $\sim 7 \times 10^{-13}$ Hz, and in the time derivative of barycentric rotation frequency is $\sim 1.8 \times 10^{-20}$ s $^{-2}$. Whilst it is then clear that $f_{\rm em}^{(n)} (t_1)$, $\dot{f}_{\rm em}^{(n)} (t_1)$ are typically measured with very high precisions, for this work we adopt a broader uniform prior, $\pm 10^3 \epsilon^{(n)}_{a}$ either side of the true value, where $\epsilon^{(n)}_{a}$ is the error for that pulsar quoted in the Australia Telescope National Facility (ATNF) catalogue and the subscript $a = \{ f, \dot{f} \}$ identifies the error as being in $f_{\rm em}^{(n)} (t_1)$, or $\dot{f}_{\rm em}^{(n)} (t_1)$ respectively. In this way we can test how well our method performs without requiring exceptionally precise and accurate measurements of the pulsar parameters to be made \textit{a priori}. Instead we consider the pulsar parameters similarly to the GW source parameters and estimate them optimally within a consistent framework. Adopting looser priors in this way allows us to subject our method to a more stringent test and confirm that we still estimate GW parameters without highly accurate initial estimates of $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$. \newline

The pulsar distances $d^{(n)}$ are typically not as well constrained as $f_{\rm em}^{(n)} (t_1)$,$\dot{f}_{\rm em}^{(n)} (t_1)$, with uncertainties typically on the order of $\sim 10 \%$ \citep{Arzoumanian2018ApJS..235...37A,Yao2017}. However for this work, whilst the pulsar distance is used when generating synthetic data, as discussed in Section \ref{sec:parameter_estim} the pulsar distance is not used for our inference model and so we do not need to set a prior on $d$. Similarly we do not set a prior on $\gamma^{(n)}$; since $[\gamma^{(n)}]^{-1} >> T_{\rm obs}$ this means that $\gamma^{(n)}$ is effectively ``unobservable" over the decadal timescales that we are interested in. That is, for $T_{\rm obs}$ =10 years the solution of Equation \eqref{eq:spinevol} is effectively independent of the choice of $\gamma^{(n)}$ as long as the condition $[\gamma^{(n)}]^{-1} \gg T_{\rm obs}$ is satisfied. It is therefore sufficient to consider $\gamma^{(n)}$ to be known \textit{a priori} and set it at its true injected value. We briefly explored setting an uninformative prior on $\gamma$ over e.g. LogUniform($10^{-15}, 10^{-10}$) as well as setting $\gamma^{(n)}$ at some fixed value away from the true injected value (e.g. set $\gamma^{(n)} = 10^{-14} \text{s}^{-1}$ rather than $10^{-13} \text{s}^{-1}$) but the results are unchanged. The majority of pulsars in our synthetic PTA have $\sigma^{(n)} \sim 10^{-20} - 10^{-22}$ as calculated from Equations \eqref{eq:sigmap} and \eqref{eq:sigmap_f}. For these pulsars we set an uninformative broad prior of LogUniform($10^{-23}, 10^{-19}$). The single exception is PSR J1939+2134 which has a particularly large $\dot{f}_{\rm p}^{(n)}$ compared to the other pulsars in the array and so $\sigma \sim 10^{-16}$. For our artificial dataset we set $\sigma = 10^{21}$ for this pulsar in order to bring the process noise for all pulsars into a consistent range that can be described by a constrained prior. \newline

Since we are not setting priors on $\gamma$ or $d$ this reduces the dimensionality of the parameter space to $7 + 3N$. We use the notation $\boldsymbol{\theta}_{\rm psr, reduced}$ to refer to the reduced parameter space. Explicitly, 
\begin{equation}
	\boldsymbol{\theta}_{\rm psr, reduced} = \left \{ f_{\rm em}^{(n)}(t_1),\dot{f}_{\rm em}^{(n)}(t_1),\sigma^{(n)}\right\}_{1\leq n \leq N} \ .
\end{equation}
All the injected static parameters and their corresponding priors for this representative example are summarised in Table \ref{tab:parameters_and_priors}.
\begin{table}
	\centering
	\resizebox{\columnwidth}{!}{%
	\renewcommand{\arraystretch}{1.5} % Default value: 1
	\begin{tabular}{lccll}
		\toprule
		&Parameter & Injected Values & Units & Prior  \\
		\hline
		\multirow{7}{2mm}{$\boldsymbol{\theta}_{\rm gw}$} & $\Omega$       & $5 \times 10^{-7}$ & Hz & LogUniform($10^{-9}$, $10^{-5}$) \\
	  & $\alpha$          & $1.0$  & radians & Uniform($0, 2 \pi $)\\
	  & $\delta$              & $1.0$  & radians & Uniform($-\pi/2, \pi/2$) \\
	  & $\psi$              & $2.50$ & radians & Uniform($0, 2 \pi $) \\
	  & $\Phi_0$          & $0.20$ & radians & Uniform($0, 2 \pi $) \\
	  & $h_0$            & $10^{-12}$ & - & LogUniform($10^{-15}$, $10^{-9}$) \\
	  & $\iota$             & $1.0$ & radians & Uniform($0, \pi$) \\ 
		\hline
		\multirow{5}{2mm}{$\boldsymbol{\theta}_{\rm psr}$} & $f_{\rm em}^{(n)} (t_1)$       & $f_{\rm ATNF}^{(n)}$ & Hz & Uniform$\left( f_{\rm ATNF}^{(n)} - 10^3 \epsilon^{(n)}_{f}, f_{\rm ATNF}^{(n)} + 10^3 \epsilon^{(n)}_{f} \right)$ \\
		& $\dot{f}_{\rm em}^{(n)} (t_1)$       & $\dot{f}_{\rm ATNF}^{(n)}$ & s$^{-2}$ & Uniform$\left( \dot{f}_{\rm ATNF}^{(n)} - 10^3 \epsilon^{(n)}_{\dot{f}}, \dot{f}_{\rm ATNF}^{(n)} + 10^3 \epsilon^{(n)}_{\dot{f}} \right)$ \\
		&  $d^{(n)}$       &$d_{\rm ATNF}^{(n)}$  & m & - \\
		& $\sigma^{(n)}$              & $\sigma_{\rm sc}^{(n)}$ & $s^{-5/2}$ & LogUniform($10^{-23}, 10^{-19}$) \\
		& $\gamma$              & $10^{-13}$ & s$^{-1}$ & - \\
		\bottomrule
	\end{tabular}
}
	\caption{Summary of injected static parameters used for generating synthetic data in the representative example of Section \ref{sec:rep_example}, along with the choice of prior used for Bayesian inference on each parameter. The subscript ``ATNF" denotes values which have been obtained from the ATNF pulsar catalogue as described in Section \ref{sec:synt_pta}. The subscript ``sc" indicates that the injected value has been calculated using the empirical timing model, Equations \eqref{eq:sigmap}, \eqref{eq:sigmap_f} from \protect \cite{Shannon2010}.}
	\label{tab:parameters_and_priors}
\end{table}

@@

\begin{figure*}
	\includegraphics[width=0.8\textwidth, height =0.8\textwidth ]{images/representative_example_v2_GW}
	\caption{Posterior distribution for the GW source parameters $\boldsymbol{\theta}_{\rm gw}$ for a single realisation of the system noise. The vertical orange lines indicate the true injected values, c.f. Table \ref{tab:parameters_and_priors}. The contours in the 2D histograms denote the (0.5, 1, 1.5, 2)-$\sigma$ levels. We are able to accurately estimate each parameter of interest. $\Omega_{\rm nHz} = \Omega \times 10^{9}$ Hz and $h_{0, \times 10^{12}} = h_0 \times 10^{12}$.}
	\label{fig:corner_plot_1}
\end{figure*}

@@

\subsubsection{Detection} \label{sec:detection}
We frame the problem of claiming a detection of a GW in the noisy data in terms of a Bayesian model selection procedure as described in Section \ref{sec:model_selection}. $\mathcal{M}_1$ is the Earth-terms only model, i.e. Equation \eqref{eq:measuremen_earth}. The Bayes factor, $\beta$, as defined in Equation \eqref{eq:bayes} is presented in Fig \ref{fig:bayes} for our representative system, where we now vary the magnitude of the GW strain, $h_0$. The noise processes in the synthetic data are identical realisations for each value of $h_0$; the only change in the data is the change in $h_0$. We can see that there is a general log-linear relationship between $\beta$ and $h_0$ where the GW source is easily detectable with decisive evidence ($\beta \sim 100$) for $h_0 \gtrapprox 10^{-14}$. We take $\beta = 10$ as a tolerance cut-off above which we can accept $\mathcal{M}_1$ and claim a detection of the GW. For this system, given this realisation of the noise, the minimal detectable strain is $\sim 3 \times 10^{-15}$. This minimal detectable strain is of course particular to this system and will be influenced in general by e.g. $T_{\rm obs}$, $\sigma_{\rm m}$ as well as the parameters of the GW source such as its location on the sky. This will be explored in more detail in Section \ref{sec:parameter_space}.

The small perturbations in the general log-linear $\beta - h_0$ trend arise due to noise in the nested sampling method. For these points the sampler converged sub-optimally for $\mathcal{M}_1$ and so the evidence $\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1)$ is lower than would be expected. This is purely a noise artefact; if one recreates the $\beta - h_0$ curve by rerunning the nested sampler these perturbations appear at different values of $h_0$, or else are not present at all. Increasing the number of live points $n_{\rm live}$ used by the nested sampler also dampens these perturbations and smooths the $\beta - h_0$ curve, since as discussed in Section \ref{sec:nested_sampling} the uncertainties in $\mathcal{Z}$ scale as $\mathcal{O}(1/ \sqrt{n_{\rm live}})$.

Below $\beta = 10$ this noise in the sampler starts to dominate and the hierarchical relationship between $\mathcal{M}_0$ and  $\mathcal{M}_1$ fails. That is, the requirement that $\mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_1) > \mathcal{Z}(\boldsymbol{Y} | \mathcal{M}_0)$ no longer holds. This is simply due to the fact that at these small strains the system is completely noise dominated and the two models $\mathcal{M}_1, \mathcal{M}_0$ cannot be separated. 
\begin{figure}
	\includegraphics[width=\columnwidth]{images/PaperBayesRatioPlotn1000_stacked}
	\caption{Bayes ratio $\beta$ between the competing models $\mathcal{M}_1$ (GW present in data) and $\mathcal{M}_0$ (GW not present in data) at different GW strain magnitudes $,h_0$, for the representative example summarised in Table \ref{tab:parameters_and_priors}. The horizontal grey dashed line labels the detection tolerance cut-off of $\beta = 10$. The minimal detectable strain, below which $\beta < 10$, is $\sim 3 \times 10^{-15}$. \textcolor{red}{TK: probably need to regenerate this figure to get rid of the "bumps" which are just artefacts from a poor sampler convergence at that point}}
	\label{fig:bayes}
\end{figure}

\subsubsection{Multiple noise realisations} \label{sec:multiple_noise}
\begin{figure*}
	\includegraphics[width=0.8\textwidth, height =0.8\textwidth ]{images/stacked_GW_plot_1000}
	\caption{As Figure \ref{fig:corner_plot_1} but for 9 realisations of the noise processes. The inferred posterior distributions show strong agreement between the different noise realisations for the majority of static parameters. The exceptions are $\iota$ and $h$ which exhibit a larger degree of variance.} 
	\label{fig:corner_plot_2}
\end{figure*}
\begin{figure*}
	\centering
	\includegraphics[width=\textwidth]{images/compare_iotas_lines2}
	\caption{1D marginalized posteriors for $\iota$ over 9 realisations of the noise where $h$ is (a) an unknown free parameter (left panel) and (b) fixed at its true value $= 10^{-12}$. The left hand panel is equivalent to the marginalized posteriors for $\iota$ shown in Figure \ref{fig:corner_plot_2} (4th row, 4th column). The thick orange line labels the true injected value. Fixing $h$ reduces the variance between different noise realisations but results in a clear bias between the true and inferred values of $\iota$.}
	\label{fig:iota}
\end{figure*}
%\begin{figure*}
%	\setkeys{Gin}{width=\linewidth}   
%	
%	\begin{subfigure}[b]{0.3\textwidth}
%		\includegraphics[width=\textwidth]{images/distribution_omega_gw}
%		\caption{$\Omega$}
%	\end{subfigure}
%	\hfill
%	\begin{subfigure}[b]{0.3\textwidth}
%		\includegraphics[width=\textwidth]{images/distribution_phi0_gw}
%		\caption{$\Phi_0$}
%	\end{subfigure}
%	\hfill	
%	\begin{subfigure}[b]{0.3\textwidth}
%		\includegraphics[width=\textwidth]{images/distribution_psi_gw}
%		\caption{$\psi$}
%	\end{subfigure}
%	\medskip
%	
%	\begin{subfigure}[b]{0.3\textwidth}
%		\includegraphics[width=\textwidth]{images/distribution_iota_gw}
%		\caption{$\iota$}
%	\end{subfigure}
%	\hfill
%	\begin{subfigure}[b]{0.3\textwidth}
%		\includegraphics[width=\textwidth]{images/distribution_delta_gw}
%		\caption{$\delta$}
%	\end{subfigure}
%	\hfill	
%	\begin{subfigure}[b]{0.3\textwidth}
%		\includegraphics[width=\textwidth]{images/distribution_alpha_gw}
%		\caption{$\alpha$}
%	\end{subfigure}
%	\medskip
%	
%	
%	\medskip
%	\begin{subfigure}[b]{0.33\textwidth}
%		\includegraphics{images/distribution_h}
%		\caption{$h$}
%	\end{subfigure}
%	
%	\caption{Distribution in the median values of the 1D marginalized posteriors inferred by nested sampling for each static parameters of $\boldsymbol{\theta}_{\rm gw}$ over 1000 noise realisations. The orange dashed line labels the true injection value and the red dashed line labels the median of the distribution. \textcolor{red}{TK: these plots are ugly. Is there a better way to show this?}}\label{fig:median_distriubutins}
%\end{figure*}

The above discussion was for a single realisation of the noise processes. It is important to confirm that our method is robust against different realisations of the noise and that the specific realisation of the noisy data used in Sections \ref{sec:parameter_estim}, \ref{sec:detection} is not particularly advantageous or convenient for our method. To this end we take our representative example from Table \ref{tab:parameters_and_priors} and generate 1000 realisations of the process noise and measurement noise. For each realisation of the data we can then independently estimate the static parameters, $\boldsymbol{\theta}$. In Figure \ref{fig:corner_plot_2} we show the results for the  estimated values of $\boldsymbol{\theta}_{\rm gw}$ for 9 realisations of the noise. This figure is analogous to the single noise realisation of Figure \ref{fig:corner_plot_1}. We plot only 9 realisations of the noise rather than the full set of 1000 realisations so as to not overcrowd the figure. As in Section \ref{sec:parameter_estim} the estimates of $\boldsymbol{\theta}_{\rm psr}$ are recovered unambiguously across the 1000 noise realisations and for brevity we do not show the results here. \newline

@@

For each realisation of the noise we infer a marginalised 1-D posterior for each parameter.

High SNR makes the posteriors effectively point estimates. Therefore maybe credible intervals dont make as much sense

%
%. In turn, each posterior distribution has a median value. We plot the distribution of these medians over the 1000 noise realisations for each parameter of $\boldsymbol{\theta}_{\rm gw}$ in Figure \ref{fig:median_distriubutins}. For each of these distributions we can also calculate the median value (i.e. the median of the medians) and compare it with the true injection value. This is shown by the red and orange dashed lines respectively in the Figure. Analogous to the results we saw in Figure \ref{fig:corner_plot_2}, the distributions of the median values of the parameters $\Omega, \Phi_0, \psi, \delta$ and $\alpha$ are very narrow, with a generally small discrepancy between the inferred value and the injected value. Conversely, the distributions for $\iota$ and $h_0$ are much broader and exhibit a bias of $\sim 0.18$ radians and $0.17 \times 10^{-12}$ respectively. This agrees with the results we discussed for the 9 noise realisations in Figure \ref{fig:corner_plot_2} where a bias in $\iota$ and $h_0$ was also observed. In addition to  $\iota$ and $h_0$  exhibiting a strong bias, the parameters $\psi$ and $\alpha$ also exhibit a small bias of magnitude $\sim 0.1$ and $\sim 0.04$ radians respectively. This bias is again a result of dropping the pulsar terms from the measurement equation and will be discussed in Section \ref{sec:bias}.

@@

%%%%%%%%%%%%%%%%Scratch space

%Taking a tangible example, for pulsar J0023+0923 ATNF returns a frequency $f_{\rm em}^{(n)} (t_1) \sim 327.8$ Hz with an error $\epsilon \sim 4 \times 10^{13}$ Hz. 	The prior on $f_{\rm em}^{(n)} (t_1) \sim 327.8$ for this pulsar is then Uniform$(327.8 - 4 \times 10^{10},327.8 + 4 \times 10^{10} )$
